Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-04-27 doc/invt.tex
update is2dos.ma

\begin{abstract}
  We present a method of computing the optimal schedule
  of the updating magnitude
  for a class of free energy calculation methods including
  the Wang-Landau (WL) algorithm and metadynamics.
  %
  These methods rely on adaptive construction of
  a bias potential that offsets
  the potential of mean force by histogram-based adaptive updates.
  %
  The updating magnitude should decrease over time
  to reduce the noise in the bias potential,
  and one may speed up the convergence by choosing an optimal schedule
  of the updating magnitude.
  %
  %Here, we will show that the optimal schedule
  %can be derived
  %from a mechanical analogy, in which
  %the schedule corresponds to
  %the velocity of a free particle with a position-dependent mass
  %and the final error serves as the action.
  %%
  %Therefore, the optimal schedule follows from
  %the equation of motion of the particle.
  %
  We will show that
  the optimal schedule depends on the updating scheme
  (such as the single-bin scheme commonly used in WL
  or the Gaussian one popular in metadynamics),
  and the initial updating magnitude is ideally half of
  the previous equilibrium value.
  %
  Further,
  the single-bin updating scheme
  belongs to a class of bandpass updating schemes
  that are optimal for asymptotic convergence,
  and the optimal schedule of these schemes
  is the inverse number of updating steps with a proper shift,
  as given by the $1/t$ prescription.
  %
\end{abstract}

>>>>

\begin{abstract}
  We present a method of computing the optimal schedule
  of the updating magnitude
  for a class of free energy calculation methods including
  the Wang-Landau (WL) algorithm and metadynamics.
  %
  These methods rely on adaptive construction of
  a bias potential that offsets
  the potential of mean force by histogram-based adaptive updates.
  %
  The updating magnitude should decrease over time
  to reduce the noise in the bias potential,
  and one may speed up the convergence by choosing an optimal schedule
  of the updating magnitude.
  %
  %Here, we will show that the optimal schedule
  %can be derived
  %from a mechanical analogy, in which
  %the schedule corresponds to
  %the velocity of a free particle with a position-dependent mass
  %and the final error serves as the action.
  %%
  %Therefore, the optimal schedule follows from
  %the equation of motion of the particle.
  %
  We will show that
  the optimal schedule depends on the updating scheme
  (such as the single-bin and Gaussian schemes commonly used in WL and metadynamics),
  and the initial updating magnitude is ideally half of
  the previous equilibrium value.
  %
  Further,
  the single-bin updating scheme
  belongs to a class of bandpass updating schemes
  that are optimal for asymptotic convergence,
  and the optimal schedule of these schemes
  is the inverse number of updating steps with a proper shift,
  as given by the $1/t$ prescription.
  %
\end{abstract}
